\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-06-09 doc/paper.tex
doc/paper.tex: minor changes
Command line is not an output channel but an input.

\fs can be downloaded as a gzipped tar file from
\url{http://flexiblesusy.hepforge.org}.  To download and install
version 1.0.0 run:
%
\begin{lstlisting}[language=bash]
$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-1.0.0.tar.gz
$ tar -xf FlexibleSUSY-1.0.0.tar.gz
$ cd FlexibleSUSY-1.0.0
\end{lstlisting}%% $
%
A CMSSM spectrum generator can be created with the following three
commands:
%
\begin{lstlisting}[language=bash]
$ ./createmodel --name=MSSM
$ ./configure --with-models=MSSM
$ make
\end{lstlisting}%% $
%
The first command creates the model directory \code{models/MSSM/}
together with a CMSSM model file.  The \code{configure} script checks
the system requirements and creates the \code{Makefile}.  See
\code{./configure --help} for more options.  Executing \code{make}
will start \mathematica to generate the spectrum generator and compile
it.  The resulting executable can be run like this:
%
\begin{lstlisting}[language=bash]
$ models/MSSM/run_MSSM.x --slha-input-file=model_files/MSSM/LesHouches.in.MSSM
\end{lstlisting}%% $
%
When executed, the spectrum generator tries to find a set of \DRbar\
model parameters consistent with all CMSSM boundary conditions for the
parameter point given in the SLHA input file
\code{model_files/MSSM/LesHouches.in.MSSM}.  Afterwards, the pole mass
spectrum and mixing matrices are calculated and written to the command
line in SLHA format \cite{Skands:2003cj,Allanach:2008qq}.  For the
parameter point given in the above example the calculated pole mass
spectrum reads
%
\begin{lstlisting}
Block MASS
   1000021     1.15236966E+03   # Glu
   1000024     3.85774334E+02   # Cha_1
   1000037     6.50460073E+02   # Cha_2
        25     1.14766149E+02   # hh_1
        35     7.06792640E+02   # hh_2
        37     7.11388516E+02   # Hpm_2
        36     7.06523105E+02   # Ah_2
   1000012     3.51856376E+02   # Sv_1
   1000014     3.53042556E+02   # Sv_2
   1000016     3.53046504E+02   # Sv_3
   1000022     2.03889780E+02   # Chi_1
   1000023     3.85760714E+02   # Chi_2
   1000025     6.36544884E+02   # Chi_3
   1000035     6.50133768E+02   # Chi_4
   1000001     9.66656018E+02   # Sd_1
   1000003     1.00983181E+03   # Sd_2
   1000005     1.01651873E+03   # Sd_3
   2000001     1.01653005E+03   # Sd_4
   2000003     1.06089534E+03   # Sd_5
   2000005     1.06090238E+03   # Sd_6
   1000011     2.22570305E+02   # Se_1
   1000013     2.29864536E+02   # Se_2
   1000015     2.29888846E+02   # Se_3
   2000011     3.61946671E+02   # Se_4
   2000013     3.61950866E+02   # Se_5
   2000015     3.63136031E+02   # Se_6
   1000002     8.09787818E+02   # Su_1
   1000004     1.01454197E+03   # Su_2
   1000006     1.01981109E+03   # Su_3
   2000002     1.02015269E+03   # Su_4
   2000004     1.05807759E+03   # Su_5
   2000006     1.05808168E+03   # Su_6
\end{lstlisting}

@@

\item The function \code{FSFindRoot[parameters, functions]} can be
  given, where \code{parameters} is a list of model parameters and
  \code{functions} is a list of functions of these parameters.
  \code{FSFindRoot[parameters, functions]} will numerical vary the
  \code{parameters} until the \code{functions} are zero.  Example:
  %
  \begin{lstlisting}
FSFindRoot[{vd,vu},
           {SM[MZ] - Pole[M[VZ]], SM[MH] - Pole[M[hh[1]]]}]
  \end{lstlisting}
  %
  Here, the parameters \code{vu} and \code{vd} are varied until the
  vector-like function
  %
  \begin{align}
    f(v_d,v_u) =
    \begin{pmatrix}
      \texttt{SM[MZ]} - m_Z^\pole \\
      \texttt{SM[MH]} - m_{h_1}^\pole
    \end{pmatrix}
  \end{align}
  %
  is zero.
\end{itemize}
%
Finally, the user can set an initial guess for the model parameters at
the low- and high-scale using the variables
\code{InitialGuessAtLowScale} and \code{InitialGuessAtHighScale},
respectively.  The gauge couplings will be guessed automatically at
the low-scale from the known Standard Model parameters.

@@

\fs can create the helper function \code{get_lsp()}, which finds the
lightest supersymmetric particle (LSP).  To have this function be
created the model file variable \code{PotentialLSPParticles} must be
set to a list of SUSY particles which are potential LSPs.  In the
model file example above, the particles \code{Chi}, \code{Cha},
\code{Glu}, \code{Sv}, \code{Su}, \code{Sd}, \code{Se} (neutralino,
chargino, gluino, sneutrino, up-type squark, down-type squark,
selectron) are considered to be LSPs.

@@

The (physical) pole mass spectrum can now be calculated.  \fs uses the
full one-loop self-energies and tree-level mass matrices obtained from
SARAH to calculate the pole masses, which means finding the values $p$
that solve the equation
%
\begin{align}
  0 = \det\left[p^2\unity - m_{f,1L}(p^2)\right].
  \label{eq:pole-mass-def}
\end{align}
%
Here the one-loop mass matrix $m_{f,1L}(p^2)$ for field $f$ is given
in terms of the tree-level mass matrix $m_f$ and the self-energy
$\Sigma_f(p^2)$ as
%
\begin{align}
  &\text{scalars } \phi: &
  m_{\phi,1L}(p^2) &= m_{\phi} - \Sigma_\phi(p^2), \\
  &\text{Majorana fermions } \chi: &
  m_{\chi,1L}(p^2) &= m_{\chi} - \frac{1}{2}\Big[
    \Sigma_\chi^S(p^2) + \Sigma_\chi^{S,T}(p^2)
    + \Big( \Sigma_\chi^{L,T}(p^2) + \Sigma_\chi^R(p^2) \Big) m_{\chi} \notag \\
    &&&\phantom{= m_{\chi} - \frac{1}{2}\Big[}
    + m_{\chi} \Big( \Sigma_\chi^L(p^2) + \Sigma_\chi^{R,T}(p^2) \Big)
  \Big], \\
  &\text{Dirac fermions } \psi: &
  m_{\psi,1L}(p^2) &= m_{\psi}
  - \Sigma_\psi^S(p^2)
  - \Sigma_\psi^R(p^2) m_{\psi}
  - m_{\psi} \Sigma_\psi^L(p^2) .
\end{align}
%
Eq.~\eqref{eq:pole-mass-def} can be solved by diagonalizing the
one-loop mass matrix $m_{f,1L}(p^2)$.  However, since $m_{f,1L}(p^2)$
depends on the momentum $p$, an iteration over $p$ must be performed.
Since this iteration can be very time consuming for large field
multiplets, \fs provides two approximative procedures with a faster
run-time.  The used procedure can be set in the model file for each
field.  These approximations work as follows:
%
\begin{itemize}
\item \code{LowPoleMassPrecision}: This option provides the
  lowest precision but is also the fastest one.  Here the one-loop
  mass matrix $m_{f,1L}^\text{low}$ is calculated exactly once as
%
  \begin{align}
    \forall i,j: (m_{f,1L}^\text{low})_{ij} = (m_{f,1L}(p^2 = m_{f_i}
    m_{f_j}))_{ij} ,
  \end{align}
%
  where $m_{f_i}$ is the $i$th mass eigenvalue of the tree-level mass
  matrix $m_f$.  Afterwards, $m_{f,1L}^\text{low}$ is diagonalized and
  the eigenvalues are interpreted as pole masses $m_{f_i}^\pole$.
  This method neglects which are formally of two-loop order.  This
  method is imprecise if the self-energy corrections to the mass
  matrix are large or the tree-level mass spectrum of the multiplet is
  very split.

>>>>

\fs can be downloaded as a gzipped tar file from
\url{http://flexiblesusy.hepforge.org}.  To download and install
version 1.0.0 run:
%
\begin{lstlisting}[language=bash]
$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-1.0.0.tar.gz
$ tar -xf FlexibleSUSY-1.0.0.tar.gz
$ cd FlexibleSUSY-1.0.0
\end{lstlisting}%% $
%
A CMSSM spectrum generator can be created with the following three
commands:
%
\begin{lstlisting}[language=bash]
$ ./createmodel --name=MSSM
$ ./configure --with-models=MSSM
$ make
\end{lstlisting}%% $
%
The first command creates the model directory \code{models/MSSM/}
together with a CMSSM model file.  The \code{configure} script checks
the system requirements and creates the \code{Makefile}.  See
\code{./configure --help} for more options.  Executing \code{make}
will start \mathematica to generate the spectrum generator and compile
it.  The resulting executable can be run like this:
%
\begin{lstlisting}[language=bash]
$ models/MSSM/run_MSSM.x --slha-input-file=model_files/MSSM/LesHouches.in.MSSM
\end{lstlisting}%% $
%
When executed, the spectrum generator tries to find a set of \DRbar\
model parameters consistent with all CMSSM boundary conditions for the
parameter point given in the SLHA input file
\code{model_files/MSSM/LesHouches.in.MSSM}.  Afterwards, the pole mass
spectrum and mixing matrices are calculated and written to the standard
output in SLHA format \cite{Skands:2003cj,Allanach:2008qq}.  For the
parameter point given in the above example the calculated pole mass
spectrum reads
%
\begin{lstlisting}
Block MASS
   1000021     1.15236966E+03   # Glu
   1000024     3.85774334E+02   # Cha_1
   1000037     6.50460073E+02   # Cha_2
        25     1.14766149E+02   # hh_1
        35     7.06792640E+02   # hh_2
        37     7.11388516E+02   # Hpm_2
        36     7.06523105E+02   # Ah_2
   1000012     3.51856376E+02   # Sv_1
   1000014     3.53042556E+02   # Sv_2
   1000016     3.53046504E+02   # Sv_3
   1000022     2.03889780E+02   # Chi_1
   1000023     3.85760714E+02   # Chi_2
   1000025     6.36544884E+02   # Chi_3
   1000035     6.50133768E+02   # Chi_4
   1000001     9.66656018E+02   # Sd_1
   1000003     1.00983181E+03   # Sd_2
   1000005     1.01651873E+03   # Sd_3
   2000001     1.01653005E+03   # Sd_4
   2000003     1.06089534E+03   # Sd_5
   2000005     1.06090238E+03   # Sd_6
   1000011     2.22570305E+02   # Se_1
   1000013     2.29864536E+02   # Se_2
   1000015     2.29888846E+02   # Se_3
   2000011     3.61946671E+02   # Se_4
   2000013     3.61950866E+02   # Se_5
   2000015     3.63136031E+02   # Se_6
   1000002     8.09787818E+02   # Su_1
   1000004     1.01454197E+03   # Su_2
   1000006     1.01981109E+03   # Su_3
   2000002     1.02015269E+03   # Su_4
   2000004     1.05807759E+03   # Su_5
   2000006     1.05808168E+03   # Su_6
\end{lstlisting}

@@

\item The function \code{FSFindRoot[parameters, functions]} can be
  given, where \code{parameters} is a list of model parameters and
  \code{functions} is a list of functions of these parameters.
  \code{FSFindRoot[parameters, functions]} will numerical vary the
  \code{parameters} until the \code{functions} are zero.  Example:
  %
  \begin{lstlisting}
FSFindRoot[{vd,vu},
           {SM[MZ] - Pole[M[VZ]], SM[MH] - Pole[M[hh[1]]]}]
  \end{lstlisting}
  %
  Here, the parameters \code{vu} and \code{vd} are varied until the
  vector-valued function
  %
  \begin{align}
    f(v_d,v_u) =
    \begin{pmatrix}
      \texttt{SM[MZ]} - m_Z^\pole \\
      \texttt{SM[MH]} - m_{h_1}^\pole
    \end{pmatrix}
  \end{align}
  %
  is zero.
\end{itemize}
%
Finally, the user can set an initial guess for the model parameters at
the low- and high-scale using the variables
\code{InitialGuessAtLowScale} and \code{InitialGuessAtHighScale},
respectively.  The gauge couplings will be guessed automatically at
the low-scale from the known Standard Model parameters.

@@

\fs can create the helper function \code{get_lsp()}, which finds the
lightest supersymmetric particle (LSP).  To have this function be
created the model file variable \code{PotentialLSPParticles} must be
set to a list of SUSY particles which are potential LSPs.  In the
model file example above, the particles \code{Chi}, \code{Cha},
\code{Glu}, \code{Sv}, \code{Su}, \code{Sd}, \code{Se} (neutralino,
chargino, gluino, sneutrino, up-type squark, down-type squark,
selectron) are considered to be LSP candidates.

@@

The (physical) pole mass spectrum can now be calculated.  \fs uses the
full one-loop self-energies and tree-level mass matrices obtained from
SARAH to calculate the pole masses, which means finding the values $p$
that solve the equation
%
\begin{align}
  0 = \det\left[p^2\unity - m_{f,1L}(p^2)\right].
  \label{eq:pole-mass-def}
\end{align}
%
Here the one-loop mass matrix $m_{f,1L}(p^2)$ for field $f$ is given
in terms of the tree-level mass matrix $m_f$ and the self-energy
$\Sigma_f(p^2)$ as
%
\begin{align}
  &\text{scalars } \phi: &
  m_{\phi,1L}(p^2) &= m_{\phi} - \Sigma_\phi(p^2), \\
  &\text{Majorana fermions } \chi: &
  m_{\chi,1L}(p^2) &= m_{\chi} - \frac{1}{2}\Big[
    \Sigma_\chi^S(p^2) + \Sigma_\chi^{S,T}(p^2)
    + \Big( \Sigma_\chi^{L,T}(p^2) + \Sigma_\chi^R(p^2) \Big) m_{\chi} \notag \\
    &&&\phantom{= m_{\chi} - \frac{1}{2}\Big[}
    + m_{\chi} \Big( \Sigma_\chi^L(p^2) + \Sigma_\chi^{R,T}(p^2) \Big)
  \Big], \\
  &\text{Dirac fermions } \psi: &
  m_{\psi,1L}(p^2) &= m_{\psi}
  - \Sigma_\psi^S(p^2)
  - \Sigma_\psi^R(p^2) m_{\psi}
  - m_{\psi} \Sigma_\psi^L(p^2) .
\end{align}
%
Eq.~\eqref{eq:pole-mass-def} can be solved by diagonalizing the
one-loop mass matrix $m_{f,1L}(p^2)$.  However, since $m_{f,1L}(p^2)$
depends on the momentum $p$, an iteration over $p$ must be performed.
Since this iteration can be very time consuming for large field
multiplets, \fs provides two approximative procedures with a shorter
run-time.  The used procedure can be set in the model file for each
field.  These approximations work as follows:
%
\begin{itemize}
\item \code{LowPoleMassPrecision}: This option provides the
  lowest precision but is also the fastest one.  Here the one-loop
  mass matrix $m_{f,1L}^\text{low}$ is calculated exactly once as
%
  \begin{align}
    \forall i,j: (m_{f,1L}^\text{low})_{ij} = (m_{f,1L}(p^2 = m_{f_i}
    m_{f_j}))_{ij} ,
  \end{align}
%
  where $m_{f_i}$ is the $i$th mass eigenvalue of the tree-level mass
  matrix $m_f$.  Afterwards, $m_{f,1L}^\text{low}$ is diagonalized and
  the eigenvalues are interpreted as pole masses $m_{f_i}^\pole$.
  This method neglects what are formally of two-loop order.  This
  method is imprecise if the self-energy corrections to the mass
  matrix are large or the tree-level mass spectrum of the multiplet is
  very split.
